\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{graphicx}
\usepackage{siunitx}
\usepackage[hyphens]{url}
\usepackage[hidelinks]{hyperref}
\usepackage{xcolor}

\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}

% Keep long inline code (file paths, dotted module names, etc.) from
% extending past the right margin: (a) allow LaTeX to stretch a problem
% line instead of overflowing; (b) walk the \code argument character by
% character and insert \penalty0 (= \allowbreak) after `.`, `/`, `-` so
% paths can wrap. Underscore breaks are not added because the common
% \code{X\_Y} idiom passes a control-sequence token, not a char.
\setlength{\emergencystretch}{3em}
\makeatletter
\def\dj@codescan#1{%
  \ifx\dj@codeend#1\else
    #1%
    \ifx#1.\penalty0\fi
    \ifx#1/\penalty0\fi
    \ifx#1-\penalty0\fi
    \expandafter\dj@codescan
  \fi}
\def\dj@codeend{\dj@codeend}
\newcommand{\code}[1]{\texttt{\dj@codescan#1\dj@codeend}}
\makeatother
\newcommand{\Msun}{M_\odot}
\newcommand{\kpc}{\mathrm{kpc}}
\newcommand{\kms}{\mathrm{km\,s^{-1}}}

\title{DwarfJeans Analysis Pipeline}
\author{Kailash Raman}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
We present a spherical Jeans modelling pipeline for Milky Way satellite
dwarf galaxies.  Per-star line-of-sight velocities are ingested via two
paths: a monolithic Geha et~al.\ (2026) catalog (Path~A) and a
collection of per-paper adapters that stage data from individual
published tables (Path~B), together covering 39 galaxies in a
registry-driven schema stored in \code{data/registry/galaxies.ecsv}.
The physical model couples an NFW dark-matter halo (scale radius
$r_s$, scale density $\rho_s$) with a Plummer stellar tracer and a
constant velocity anisotropy $\beta$, solved via the projected
spherical Jeans equation.  Inference proceeds with \textsc{dynesty}
nested sampling over seven parameters --- $\bar V$, $\log_{10}\rho_s$,
$\log_{10} r_s$, a symmetrized anisotropy $\tilde\beta$, and three
Gaussian nuisances marginalising over distance, ellipticity, and
projected half-light radius --- under a user-selectable prior on
$(\ln\rho_s,\ln r_s)$: uniform, log-uniform, or conditional Jeffreys.
Posterior samples are propagated to $\sigma_\mathrm{los}(R_{1/2,\mathrm{2D}})$,
$M(r_{1/2})$, and tidally-truncated NFW $J$- and $D$-factors at four
reporting angles ($0.1^\circ$, $0.2^\circ$, $0.5^\circ$, and $\alpha_c$
or $\alpha_c/2$ respectively).  Calibration uses a single deterministic
Asimov realisation and a 15-realisation Monte Carlo mock-recovery study
on synthetic UFD galaxies, following the gold-standard convention of
generating synthetic input at known truth and reporting bias and
dispersion across realisations.
\end{abstract}

\noindent\emph{This document describes the pipeline as implemented
in code. Repository layout, module boundaries, and migration history
live in \code{ARCHITECTURE.md} at the repository root; the working
design plan, P\&S~2018 deviation log, and per-stage extended treatments
(including derivations) live in \code{docs/plan/} (overview entry
point: \code{pipeline\_overview.md}).  Each section below opens with
a pointer to the relevant plan document so a reader can drill into
the derivation, the convention, or the historical rationale without
chasing references through the prose.  This writeup does not duplicate
either the architecture file or the plan documents.}

\tableofcontents

% ===================================================================
\section{Stage 0a: Galaxy registry}
\label{sec:stage0a}

The full provenance log for the LVDB pin, the staging conventions, and
the per-source rationale lives in \code{docs/plan/data\_sources.md} (LVDB
section); this section summarizes the as-implemented build.

\subsection{Data source and study sample}

The galaxy registry is built by \code{src/dwarfjeans/ingest/stage0a\_registry.py}
from the Local Volume Database (LVDB) pinned to version~1.0.5 (Zenodo DOI
\href{https://doi.org/10.5281/zenodo.15476348}{10.5281/zenodo.15476348}).
The combined catalog is staged on disk at
\code{data/lvdb\_v1.0.5/comb\_all.csv}; the build script verifies a
SHA-256 checksum against \code{data/lvdb\_v1.0.5/checksums.sha256}
and aborts on mismatch.  No network access occurs at run time.

The 39-galaxy study sample is enumerated in
\code{src/dwarfjeans/ingest/config/study\_sample.yaml}.  Each entry
carries four fields: \code{lvdb\_key} (the LVDB \code{key} string used
to select the row from \code{comb\_all.csv}), \code{study\_name}
(the human-readable label used in plots and tables), \code{path}
(\code{A} for Geha et~al.\ 2026 DEIMOS data or \code{B} for the LVDB
\code{ref\_vlos} paper; used by Stage~0b), and \code{geha\_galaxy}
(Path~A only: the abbreviation matching the \code{system\_name} column of
\code{data/geha2026/table5A\_20260110.csv}).  The study sample
contains 22 Path~A galaxies and 17 Path~B galaxies.

After loading \code{comb\_all.csv}, the build script applies three
mandatory filters.  First, every study-sample \code{lvdb\_key} must
resolve to exactly one row in \code{comb\_all.csv}; an unmatched key
is a hard error.  Second, all 39 rows must satisfy
\code{confirmed\_dwarf}~$= 1$.  Third, each row must have non-null
values for \code{rhalf}, \code{distance\_modulus}, and
\code{apparent\_magnitude\_v}.  Host-chain resolution verifies that
every galaxy traces transitively to the Milky Way; the immediate host
is read from the LVDB \code{host} column and the chain is followed
through the full \code{comb\_all.csv} so that subhalo hosts such as
\code{lmc} $\to$ \code{mw} resolve correctly.  The resolved chain is
stored as a \code{->}-delimited string (e.g.\ \code{lmc->mw} for the
four Carina-region LMC satellites).

\subsection{Coordinate and structural transformations}

Distance in kpc is derived from the distance modulus $\mu$ as
\begin{equation}
  d \;[\kpc] = 10^{(\mu - 10)/5}.
  \label{eq:dist}
\end{equation}
The LVDB \code{rhalf} column gives the 2D major-axis half-light
radius $R_{1/2,\mathrm{maj}}$ in arcminutes; it is converted to
physical units as
\begin{equation}
  R_{1/2,\mathrm{maj}} \;[\mathrm{pc}]
  = \theta\,[\mathrm{arcmin}]\times \frac{\pi}{180\times 60}
    \times d\,[\kpc]\times 1000.
  \label{eq:rhalf_major}
\end{equation}
The azimuthally averaged (sphericalised) 2D half-light radius, which
is the quantity entering the Jeans likelihood as $r_{1/2}$ in
P\&S~2018 eq.~8, is
\begin{equation}
  r_{1/2} \;[\mathrm{pc}]
  = R_{1/2,\mathrm{maj}}\,\sqrt{1-\varepsilon},
  \label{eq:rhalf_sph}
\end{equation}
where $\varepsilon = 1 - b/a$ is the LVDB ellipticity.  The LVDB
pre-derived columns \code{rhalf\_physical} and
\code{rhalf\_sph\_physical} are never consumed as inputs; they serve
only as cross-checks against equations~\eqref{eq:rhalf_major}--\eqref{eq:rhalf_sph}
to tolerances of 3\% (major-axis) and 7\% (sphericalised).  The wider
sphericalised tolerance reflects Jensen's inequality: the LVDB values
are MC medians over the joint $(\code{rhalf}, \varepsilon, d)$ error
distributions, which the concave $\sqrt{1-\varepsilon}$ factor
systematically lowers relative to our deterministic central value.
Empirically the worst observed offsets are $1.6\%$ for the major-axis
check (Tucana~V) and $5.5\%$ for the sphericalised check (Leo~V); a
failure on either check aborts the build.

\subsection{Ellipticity missing-data convention}

When a galaxy has no measured ellipticity in LVDB~v1.0.5
(\code{ellipticity} null), the registry records
\code{ellipticity\_missing = True} and the deterministic computation
uses $\varepsilon = 0$ in equation~\eqref{eq:rhalf_sph}; this is the
correct central value under the uniform prior $U(0, 0.5)$ that
Stage~2 substitutes for the missing measurement.  The
\code{ellipticity\_ul} column preserves any upper-limit value
published by LVDB (e.g.\ Crater~II has \code{ellipticity\_ul} $=0.10$)
for diagnostics; it does not enter any Stage~0a computation.

\subsection{Spatial model and Plummer scale radius}

LVDB~v1.0.5 does not provide a per-galaxy photometric profile flag.
The spatial model is read from a hand-curated override file,
\code{src/dwarfjeans/ingest/config/spatial\_model\_overrides.yaml},
which specifies a default (currently \code{plummer}, matching the
P\&S~2018 convention for ultra-faint dwarfs) and a per-galaxy
\code{overrides} dict; the overrides dict is currently empty so all
39 galaxies use the Plummer model.  For \code{spatial\_model = "plummer"},
the Plummer scale radius equals the sphericalised half-light radius:
$r_p = r_{1/2}$.  An override to \code{"exponential"} would require
the conversion $r_p = 1.68\,r_\mathrm{exp}$ (P\&S~2018 \S3 footnote~4)
and is currently guarded by \code{NotImplementedError} because LVDB
does not expose a per-galaxy $r_\mathrm{exp}$ value.

\subsection{Output schema}

The registry is written to \code{data/registry/galaxies.ecsv} as an
Astropy ECSV~v1.0 file with units in column metadata.  The 39 rows
are ordered to match \code{study\_sample.yaml}.  A human-readable
build log is written in parallel to
\code{data/registry/build\_log.txt}.  Columns group by purpose as
follows; column-by-column provenance is in
\code{docs/plan/data\_sources.md}.

\begin{itemize}
  \item \textbf{Identification}: \code{lvdb\_key}, \code{study\_name},
        \code{path} (A/B), \code{geha\_galaxy} (Path~A label),
        \code{host}, \code{host\_chain}.
  \item \textbf{Sky position}: \code{ra\_deg}, \code{dec\_deg} [deg].
  \item \textbf{Structural inputs}: \code{rhalf\_arcmin} (with $\pm$
        columns), \code{ellipticity} (with $\pm$ columns and
        \code{ellipticity\_ul}), \code{ellipticity\_missing} flag,
        \code{spatial\_model}.
  \item \textbf{Derived radii}: \code{rhalf\_major\_pc}
        (eq.~\ref{eq:rhalf_major}), \code{r\_half\_2d\_pc}
        (eq.~\ref{eq:rhalf_sph}), \code{plummer\_radius\_pc}
        ($= r_{1/2}$ under the current default).
  \item \textbf{Distance}: \code{distance\_modulus} (with $\pm$
        columns), \code{distance\_kpc} (eq.~\ref{eq:dist}).
  \item \textbf{Kinematics}: \code{vlos\_systemic\_kms} (with $\pm$
        columns), \code{vlos\_sigma\_kms},
        \code{vlos\_sigma\_unresolved} flag.  The LVDB
        $\sigma_\mathrm{los}$ value is metadata only; it does not
        enter the Jeans likelihood.
  \item \textbf{Photometry}: \code{M\_V}.
  \item \textbf{LVDB provenance}: \code{ref\_vlos},
        \code{ref\_structure} --- NASA ADS bibcodes for the LVDB
        historical kinematic and structural references.  These are
        pedigree fields and are \emph{not} guaranteed to identify the
        catalog actually consumed by our analysis.
  \item \textbf{Analysis source}: \code{data\_source\_paper} ---
        bibcode of the paper whose per-star catalog is loaded into
        \code{data/star\_catalogs/<lvdb\_key>.npz}.  For Path~A this
        is \code{Geha2026arXiv260210200G} regardless of LVDB's
        \code{ref\_vlos}; for Path~B this matches \code{ref\_vlos}
        (the adapter dispatch key).  Downstream code, writeup
        citations, and comparisons against published work should use
        this column, not \code{ref\_vlos}.
\end{itemize}

% ===================================================================
\section{Stage 0b: Per-galaxy stellar kinematic ingest}
\label{sec:stage0b}

The two-path ingest design (Geha 2026 monolithic catalog vs.\ per-paper
adapters) and the no-cuts-at-ingest invariant are documented in
\code{docs/plan/data\_sources.md}; per-paper combiner conventions
(IVW recipes, $\sigma_\mathrm{sys}$ floors, instrument zero-point
offsets) are tabulated in \code{docs/plan/per\_paper\_combiners.md}.
This section summarizes the two paths as they are implemented.

\subsection{Path A: Geha+2026 ingest}
\label{sec:stage0b-patha}

The 22 Path~A galaxies are ingested from the homogeneous Keck/DEIMOS
spectroscopic census of Geha et~al.\ (2026, Papers~I/II).  The driver
is \code{src/dwarfjeans/ingest/stage0b\_geha.py}; it reads the
pre-staged full-precision per-star catalog
\code{data/geha2026/table5A\_20260110.csv} (Paper~I,
24,436 rows $\times$ 50 columns), filters per-galaxy slices by the
\code{system\_name} column (matched to the registry field
\code{geha\_galaxy}), and writes one
\code{data/star\_catalogs/<lvdb\_key>.npz} archive per galaxy.  Table 5A
is the authoritative source as of 2026-05-08 because it carries the
binary \code{Pmem\_novar} column (1 = member with velocity variables
already removed; 0 = non-member or variable) used by Paper~II~\S3.1 to
compute the Table~A1 N\* counts.  The legacy
\code{table3A\_20260110.csv} release exposes only a graded \code{Pmem}
and a separate \code{Var} flag; combining them as
$\mathrm{Pmem}>0.5\,\&\,\mathrm{Var}\neq 1$ does not exactly reproduce
Paper~II's N\*.

\paragraph{Output schema.}
Every Path~A archive carries the canonical columns required by the
downstream Jeans inference: \code{V} and \code{sigma\_eps}
(heliocentric line-of-sight velocity and 1$\sigma$ error in $\kms$,
mapped from \code{v} and \code{v\_err}); \code{p} (binary 0/1
membership flag, copied verbatim from \code{Pmem\_novar}); \code{R}
(projected physical radius in kpc, computed via
\code{staging.projected\_radius\_kpc}); \code{star\_id} (zero-based
contiguous integer per galaxy, with per-star uniqueness verified);
and \code{RA\_star}, \code{Dec\_star}.  Auxiliary columns retained
verbatim include \code{FeH}, \code{FeH\_err}, \code{Var}
(from \code{flag\_var}), \code{Pmem} (the graded probability, kept
for cross-checks), \code{g\_mag}, \code{r\_mag}, \code{MV},
\code{nmask}, \code{t\_exp}, \code{SN}, \code{CaT}, \code{CaTerr}.  Stage~0b is raw-data-only: membership thresholding
and aperture cuts are deferred to a downstream selection stage.

\subsection{Path B: per-paper adapters}
\label{sec:stage0b-pathb}

The remaining 17 galaxies are ingested from 15 individual publications
via the per-paper adapter system in
\code{src/dwarfjeans/ingest/stage0b\_pathb.py} dispatching to modules
under \code{src/dwarfjeans/ingest/path\_b\_adapters/}.

\paragraph{Adapter contract.}
Each adapter exposes a single function
\begin{quote}
\code{load(staged\_dir, registry\_row) -> (arrays, meta\_extra)}
\end{quote}
where \code{staged\_dir} points to the checksum-verified raw-data
directory and \code{registry\_row} carries the LVDB metadata for the
target galaxy.  The driver requires six array keys (\code{V},
\code{sigma\_eps}, \code{p}, \code{star\_id}, \code{RA\_star},
\code{Dec\_star}), all of identical length, plus a \code{meta\_extra}
dict that is merged into the archive's \code{\_meta} JSON blob.
Adapters with \code{catalog\_granularity = "per\_star"} must emit
unique \code{star\_id} values; per-epoch adapters are permitted to
repeat a \code{star\_id} across rows (one per observation).  Dispatch
is by bibkey \code{<lastname><year>} derived from the LVDB
\code{ref\_vlos} field; unknown adapters and unknown instrument tags
raise rather than falling back silently.

\paragraph{Papers and galaxies covered.}
Table~\ref{tab:pathb-adapters} lists the 15 adapter modules (covering
17 galaxies) currently in the Path~B ingest.

\begin{table}[h]
\centering
\caption{Path~B adapter modules.  ``Granularity'' denotes whether the
  adapter's output is per-star (one row per star, pre-combined
  velocity) or per-epoch (one row per observation, fed through the
  multi-epoch combiner).}
\label{tab:pathb-adapters}
\begin{tabular}{lll}
\hline
Adapter module & Galaxy / galaxies (LVDB key) & Granularity \\
\hline
\code{bruce2023}   & \code{aquarius\_2}                            & per-star  \\
\code{chiti2022}   & \code{grus\_1}                                & per-epoch \\
\code{chiti2023}   & \code{tucana\_2}                              & per-epoch \\
\code{hansen2024}  & \code{tucana\_5}                              & per-epoch \\
\code{heiger2024}  & \code{centaurus\_1}                           & per-star  \\
\code{ji2021}      & \code{antlia\_2}, \code{crater\_2}            & per-star  \\
\code{kirby2015}   & \code{pisces\_2}                              & per-star  \\
\code{koposov2015} & \code{horologium\_1}                          & per-star  \\
\code{koposov2018} & \code{hydrus\_1}                              & per-star  \\
\code{li2017}      & \code{eridanus\_2}                            & per-epoch \\
\code{li2018}      & \code{carina\_2}, \code{carina\_3}            & per-epoch \\
\code{simon2020}   & \code{tucana\_4}                              & per-epoch \\
\code{tan2025}     & \code{leo\_6}                                 & per-star  \\
\code{walker2009}  & \code{carina\_1}                              & per-star  \\
\code{walker2015}  & \code{reticulum\_2}                           & per-star  \\
\hline
\end{tabular}
\end{table}

Per-star adapters emit pre-combined velocities directly; per-epoch
adapters emit one row per spectrograph exposure, which the multi-epoch
combiner (Section~\ref{sec:stage0b-multiepoch}) reduces to per-star
rows.  Several per-epoch sources include stars flagged as binary
candidates by the source authors (e.g.\ Tuc~V-1 in Hansen+2024,
TucII-309 / TucII-078 in Chiti+2023); these flags are carried as
auxiliary columns but no rows are dropped at ingest.  Binary-aware
sample selection is deferred.

\subsection{Multi-epoch combination}
\label{sec:stage0b-multiepoch}

Per-epoch catalogs are reduced to per-star arrays in
\code{src/dwarfjeans/ingest/multi\_epoch.py}; the dispatched combiner
in \code{src/dwarfjeans/ingest/combiners/default.py} groups rows by
\code{star\_id} and applies inverse-variance weighting (IVW) under a
per-paper \code{CombinePolicy} (\code{combiners/\_\_init\_\_.py},
frozen dataclass).  Dispatch is keyed by
\code{source\_paper\_bibcode} stamped into each adapter's
\code{meta\_extra}; per-paper handlers currently cover Li+2017
(Eri~II), Li+2018 (Car~II/III), Chiti+2022 (Gru~I), Chiti+2023 (Tuc~II),
Simon+2020 (Tuc~IV), and Hansen+2024 (Tuc~V).

\paragraph{Default IVW (the ``$\sigma_\mathrm{sys}$-as-statistical'' path).}
Under the default convention, the published per-epoch error
$\sigma_{\varepsilon,i}$ is fed directly into the IVW.  For a star
with $N$ epochs,
\begin{equation}
  \bar v
  = \frac{\sum_{i=1}^{N} v_i / \sigma_{\varepsilon,i}^2}
         {\sum_{i=1}^{N} 1/\sigma_{\varepsilon,i}^2},
  \qquad
  \sigma_{\bar v}
  = \sqrt{\frac{1}{\sum_{i=1}^{N} 1/\sigma_{\varepsilon,i}^2}
          + \sigma_\mathrm{sys}^2},
  \label{eq:ivw}
\end{equation}
with $\sigma_\mathrm{sys} = 0$ for every per-paper handler under the
current convention.

\paragraph{Bias from the default convention.}
Every UFD spectroscopy paper in the current ingest reports per-epoch
errors that already include a systematic floor in quadrature,
$\sigma_\varepsilon^2 = \sigma_\mathrm{stat}^2 + \sigma_\mathrm{sys}^2$.
Feeding $\sigma_\varepsilon$ into eq.~\eqref{eq:ivw} treats
$\sigma_\mathrm{sys}$ as if it were statistical, so the IVW averages
it down by $1/\sqrt N$ along with $\sigma_\mathrm{stat}$.  The
combined uncertainty is therefore biased low by approximately
$\sigma_\mathrm{stat}/\sqrt{\sigma_\mathrm{stat}^2 + \sigma_\mathrm{sys}^2}$
--- a $\sim 10$--$30\%$ underestimate at typical $N=2$--$5$ when
$\sigma_\mathrm{sys} \sim \sigma_\mathrm{stat}$.  The variability
$\chi^2$ $p$-values are correspondingly biased high.  The bias is
below the chain-noise floor for current $\sigma_\mathrm{los}\sim$
few-$\kms$ inferences but is non-negligible for galaxies with
$\le 3$ surviving members (e.g.\ Tuc~V); see the
\code{multi\_epoch.py} module docstring and
\code{docs/plan/per\_paper\_combiners.md}.

\paragraph{Strict deconvolution (opt-in).}
Setting \code{CombinePolicy.sigma\_sys\_kms} to a positive value
routes the handler through \code{combine\_star\_strict} /
\code{ivw\_combine\_strict} in \code{multi\_epoch.py}.  Per-epoch
statistical components are recovered as
$\sigma_{\mathrm{stat},i} = \sqrt{\sigma_{\varepsilon,i}^2 - \sigma_\mathrm{sys}^2}$,
the IVW operates on $\sigma_{\mathrm{stat},i}$, and
$\sigma_\mathrm{sys}$ is re-added in quadrature post-combine:
\begin{equation}
  \bar v_\mathrm{strict}
  = \frac{\sum_{i=1}^{N} v_i / \sigma_{\mathrm{stat},i}^2}
         {\sum_{i=1}^{N} 1/\sigma_{\mathrm{stat},i}^2},
  \qquad
  \sigma_{\bar v,\,\mathrm{strict}}
  = \sqrt{\frac{1}{\sum_{i=1}^{N} 1/\sigma_{\mathrm{stat},i}^2}
          + \sigma_\mathrm{sys}^2}.
  \label{eq:ivw-strict}
\end{equation}
This is the textbook treatment of a fully correlated instrument
systematic; no per-paper handler currently activates it.

\subsection{Per-instrument zero-point offsets}
\label{sec:stage0b-zeropoint}

\code{CombinePolicy.zero\_point\_offsets\_kms} maps an instrument tag
(read from the per-epoch \code{Inst} column) to an additive shift
applied to $V$ before the IVW:
\begin{equation}
  v_i' = v_i + \delta_\mathrm{zp}(\code{Inst}_i).
  \label{eq:zeropoint}
\end{equation}
A non-empty offset table requires an \code{Inst} column; a missing
column or unknown tag raises (silent zero-offset fallback is
explicitly forbidden).  The reference instrument is defined by
assigning offset zero.  Chiti~et~al.\ (2023; Tucana~II) is the only
handler currently using non-zero offsets.  With MIKE as reference,
the \code{chiti2023} policy applies
$\delta_\mathrm{zp} = \{$MIKE$=0$, M2FS$=+2.5$, IMACS$=+2.2$, MagE$=+1.0\}\,\kms$
per the inter-instrument offsets reported in that paper's \S3.1
(byte-verified against the staged per-epoch table; see
\code{docs/plan/per\_paper\_combiners.md}).  Application shifts
individual $\bar v$ values by up to $+2.5\,\kms$ and reduces the
proxy $\sigma_\mathrm{los}$ across the 19 Tuc~II member stars from
$4.02$ to $3.88\,\kms$ ($\approx 3\%$).

\subsection{Variability flagging}
\label{sec:stage0b-varflag}

Stars with $N\ge 2$ epochs are tested for velocity variability:
\begin{equation}
  \chi^2 = \sum_{i=1}^{N} \frac{(v_i - \bar v)^2}{\sigma_i^2},
  \qquad \mathrm{dof} = N - 1,
  \label{eq:chi2var}
\end{equation}
with $\sigma_i = \sigma_{\varepsilon,i}$ under the default path and
$\sigma_i = \sigma_{\mathrm{stat},i}$ under the strict path.  The
upper-tail $p$-value is compared against
\code{CombinePolicy.p\_threshold} (framework default $0.01$): if
$p < p_\mathrm{threshold}$, the star receives \code{var\_flag = True}.
Single-epoch stars are returned with $\chi^2 = p = \mathrm{NaN}$ and
\code{var\_flag = False}; no test is performed.  The framework
default of $0.01$ matches Simon+2020 \S3.4 explicitly and is
consistent with the Li+2018 / Chiti+2023 hand-flag rules of thumb on
inter-epoch $\Delta v$.

\subsection{Per-paper choices summary}
\label{sec:stage0b-perpaper}

The canonical table of per-paper combiner settings is maintained in
\code{docs/plan/per\_paper\_combiners.md}.  All seven per-paper
handlers currently set
\code{CombinePolicy.sigma\_sys\_kms = 0.0} --- not because the source
papers omit a systematic floor, but precisely because every paper in
the per-epoch set (Li+2017, Li+2018, Chiti+2022, Chiti+2023,
Simon+2020, Hansen+2024) reports a per-instrument
$\sigma_\mathrm{sys}$ in the range $0.5$--$1.2\,\kms$ that is
\emph{already} added in quadrature into the published $e_\mathrm{RVel}$
column.  Setting \code{sigma\_sys\_kms > 0} would double-count.  All
handlers retain \code{p\_threshold = 0.01}; only \code{chiti2023}
populates a non-empty \code{zero\_point\_offsets\_kms} mapping.

\subsection{Star selection}
\label{sec:stage0b-selection}

Per-star catalogs from Stages~0b are reduced to the kinematic sample
fed to Stage~1 by \code{select\_jeans\_stars} in
\code{src/dwarfjeans/jeans/selection.py}, called automatically by
\code{prepare\_jeans\_input}.  Three cuts are applied in order:

\begin{enumerate}
  \item \textbf{Membership}: $p_i > p_\mathrm{min}$ (default
        $p_\mathrm{min} = 0.5$).  For Path~A, $p_i$ is the binary
        \code{Pmem-novar} from Paper~I (1 = member with velocity
        variables already removed, 0 otherwise).  For Path~B,
        $p_i$ is the per-paper adapter's published value.  Catalogs
        whose $p$ is uniformly~1 (already hard-cut upstream) skip
        the cut.
  \item \textbf{Radial}:
        $R_i < r_\mathrm{cut} \equiv 2 \, r_{1/2,\mathrm{maj}}\,
        \sqrt{1 - \varepsilon}\, \cdot 4/3$, the sphericalised 3D
        Plummer half-mass radius scaled by 2.  This convention is
        empirically tuned to reproduce Paper~II Table~A1 N\* across
        the Path~A sample (17/22 galaxies within $\pm 2$ stars; the
        five outliers — Leo~I, Leo~II, Sextans, Ursa~Minor, Hercules
        — differ by $\le 26$ stars and trace to \code{Pmem-novar}
        snapshot drift, not to the cut definition).  For galaxies
        with no published ellipticity (\code{ellipticity\_missing =
        True}), $\varepsilon = 0$ is used and the missing-flag is
        recorded in the audit.  The constant $4/3$ is the
        conventional round approximation to the Plummer 3D/2D ratio
        and is kept distinct from the exact $1.30477$ used by the
        derived $M_{1/2,\mathrm{3D}}$ chain (Section~\ref{sec:stage3});
        the $\sim$2.2\% discrepancy reflects the empirical, not
        analytical, basis of the cut.
  \item \textbf{Variability}: drop velocity-variable stars
        (\code{var\_flag = True} or \code{Var = 1.0}).  For Path~A
        the variability removal is already folded into
        \code{Pmem-novar} upstream, so this cut is a no-op
        (defensive belt-and-braces).
\end{enumerate}

The selection report (audited per-galaxy in \code{audit.json}) records
\code{n\_input}, \code{n\_after\_p}, \code{n\_after\_R},
\code{n\_after\_var}, \code{n\_final}, \code{rhalf\_major\_pc},
\code{ellipticity}, \code{ellipticity\_missing},
\code{rhalf\_sph\_3d\_pc}, and \code{R\_max\_kpc}.

% ===================================================================
\section{Stage 1: Spherical Jeans inference}
\label{sec:stage1}

The canonical extended treatment of the Jeans inference (motivation,
parameter parameterizations, log-space numerical scheme, sampler
configuration, MC validation plan) lives in \code{docs/plan/stage2.md}
in the working plan; the high-level role in the broader pipeline is
in \code{docs/plan/pipeline\_overview.md}.  The full algebraic
derivation of the conditional Jeffreys prior $p(\ln\rho_s,\ln r_s\mid
\beta,r_p,\mathrm{data})$ used here lives in
\code{docs/plan/jeffreys\_jeans\_derivation.md}; this section quotes
the resulting formula but not the full derivation.  Nuisance-prior
conventions (split-normal targets, the symmetric-Gaussian interim) are
in \code{docs/plan/uncertainty\_conventions.md}.  The Walker constant-$\sigma$
baseline reported alongside (Section~\ref{sec:stage1:walker}) is the
plan's Stage~1 (\code{docs/plan/stage1.md}) — it's bundled here in the
writeup because it shares the same per-galaxy driver.

\subsection{Model}
\label{sec:stage1-model}

The pipeline models each dwarf as a spherically symmetric system in
dynamical equilibrium, with an NFW dark-matter halo, a Plummer stellar
tracer, and constant velocity anisotropy.

\paragraph{NFW halo.}
The dark-matter density profile is
\begin{equation}
  \rho_\mathrm{DM}(r) = \frac{\rho_s}{(r/r_s)(1+r/r_s)^2},
  \label{eq:nfw-rho}
\end{equation}
with $\rho_s$ in $\Msun\,\kpc^{-3}$ and $r_s$ in kpc.  The enclosed
mass follows analytically:
\begin{equation}
  M(r) = 4\pi\,\rho_s\,r_s^3\;g(r/r_s),
  \qquad
  g(x) = \ln(1+x) - \frac{x}{1+x},
  \label{eq:nfw-mass}
\end{equation}
implemented in \code{src/dwarfjeans/jeans/solver.py} (\code{nfw\_M},
\code{nfw\_g}).  For numerical stability $g(x)$ is evaluated via a
Taylor series for $x < 10^{-3}$.  No tidal-radius truncation is
applied at this stage; the integration upper limit is set to
$100\max(r_s, r_p)$, well into the Plummer $r^{-5}$ tail.

\paragraph{Plummer tracer.}
The 3D stellar number density and projected surface density are
\begin{equation}
  \nu(r) = \frac{3}{4\pi r_p^3}\left(1 + \frac{r^2}{r_p^2}\right)^{-5/2},
  \qquad
  \Sigma(R) = \frac{1}{\pi r_p^2}\left(1 + \frac{R^2}{r_p^2}\right)^{-2},
  \label{eq:plummer}
\end{equation}
where $r_p$ is the 3D Plummer scale (\code{plummer\_nu},
\code{plummer\_Sigma} in \code{solver.py}).  For the Plummer profile
the projected half-light radius equals $r_p$.  In 7-parameter
nuisance mode, the physical Plummer scale is derived per draw as
\begin{equation}
  r_p = d \cdot \vartheta_{1/2}\cdot\frac{\pi}{180\times 60}\,\sqrt{1-\varepsilon}.
  \label{eq:rp_nuisance}
\end{equation}

\paragraph{Anisotropy.}
Velocity anisotropy is parametrised by the constant
$\beta \equiv 1 - \sigma_\theta^2/\sigma_r^2$, sampled through the
symmetrised variable
\begin{equation}
  \tilde\beta \in [-0.95,\,1),
  \qquad
  \beta = \frac{2\tilde\beta}{1 + \tilde\beta},
  \label{eq:beta-tilde}
\end{equation}
which maps the physical range $\beta\in(-\infty,1)$ onto a bounded
interval that is numerically convenient for nested sampling
(\code{beta\_tilde\_to\_beta} in \code{src/dwarfjeans/jeans/inference.py}).
The tangential edge $\tilde\beta = -0.95$ corresponds to
$\beta\approx -38$.

\paragraph{Jeans equation.}
The spherical Jeans equation for constant $\beta$,
\begin{equation*}
  \frac{1}{\nu}\frac{d(\nu\sigma_r^2)}{dr} + \frac{2\beta\,\sigma_r^2}{r}
  = -\frac{G M(r)}{r^2},
\end{equation*}
admits the integrating factor $r^{2\beta}$ and yields the
closed-form integral solution (Pace \& Strigari 2018, eqs.~3--7;
hereafter P\&S~2018):
\begin{equation}
  \nu(r)\,\sigma_r^2(r)
  = r^{-2\beta}
    \int_r^{\infty} s^{2\beta}\,\nu(s)\,\frac{G\,M(s)}{s^2}\,ds.
  \label{eq:jeans-J1}
\end{equation}
This is integral J1 of the module docstring.  Production evaluation
uses a log-spaced grid in $s$ with reverse cumulative log-sum-exp
(\code{nu\_sigma\_r2\_grid} in \code{solver.py}) to avoid float64
overflow at the extreme tangential edge where $2\beta \approx -76$.

\paragraph{Projection integral.}
The line-of-sight velocity dispersion at projected radius $R$ is
obtained from eq.~\eqref{eq:jeans-J1} via the constant-$\beta$
projection kernel (P\&S~2018 eq.~4):
\begin{equation}
  \Sigma(R)\,\sigma_\mathrm{los}^2(R)
  = 2\int_R^{\infty}
      \left(1 - \beta\,\frac{R^2}{r^2}\right)
      \frac{\nu(r)\,\sigma_r^2(r)\,r}{\sqrt{r^2 - R^2}}\,dr.
  \label{eq:projection-P1}
\end{equation}
The square-root singularity at $r = R$ is regularised by the
substitution $r^2 = R^2 + u^2$, giving a smooth integrand on
$u \in [0,\infty)$ that is integrated by a per-$R$ log-spaced
trapezoidal rule with an explicit $u=0$ endpoint
(\code{Sigma\_sigma\_los2\_grid} in \code{solver.py}).
The public interface returns
$\sigma_\mathrm{los}(R) = \sqrt{\Sigma\sigma_\mathrm{los}^2 / \Sigma}$
(\code{sigma\_los}).

\subsection{Likelihood}
\label{sec:stage1-likelihood}

\paragraph{Per-star Gaussian.}
Each star $i$ has a measured line-of-sight velocity $V_i$, a combined
measurement uncertainty $\sigma_{\varepsilon,i}$ (which already
incorporates the instrument systematic floor from Stage~0b), a
projected radius $R_i$, and a membership probability $p_i$.  The
total per-star variance is
\begin{equation}
  s_i^2 \equiv \sigma_\mathrm{los}^2(R_i;\,\rho_s, r_s, \beta, r_p)
              + \sigma_{\varepsilon,i}^2,
  \label{eq:stot}
\end{equation}
and the per-star log-Gaussian is
\begin{equation}
  \ln\mathcal{L}_i
  = -\tfrac{1}{2}\ln\!\bigl(2\pi s_i^2\bigr)
    - \frac{(V_i - \bar V)^2}{2\,s_i^2}.
  \label{eq:per-star-lnlike}
\end{equation}
No additional intrinsic scatter term is added here; the systematic
floor is already in $\sigma_{\varepsilon,i}$.

\paragraph{Membership-weighted pseudo-likelihood.}
Membership enters as an exponent on each per-star density (Walker
et~al.\ 2006 convention).  The total log-likelihood summed over $N$
stars is
\begin{equation}
  \ln\mathcal{L}
  = \sum_{i=1}^{N} p_i\,\ln\mathcal{L}_i
  = \sum_{i=1}^{N} p_i
    \left[-\tfrac{1}{2}\ln(2\pi s_i^2)
          -\frac{(V_i - \bar V)^2}{2\,s_i^2}\right],
  \label{eq:lnlike}
\end{equation}
implemented as \code{ll = float(np.sum(p * ln\_li))} in
\code{inference.make\_loglike} and
\code{inference.make\_loglike\_with\_nuisances}.  Stars with $p_i = 1$
(post-membership-cut samples) reduce eq.~\eqref{eq:lnlike} to an
ordinary unweighted sum; the general $p_i$-weighted form is used for
catalogs that retain probabilistic membership.  Stars below the
membership threshold $p_\mathrm{min}$ are excluded by
\code{prepare\_jeans\_input} before inference begins.  The Fisher
information of this pseudo-likelihood, used by the Jeffreys prior
(Section~\ref{sec:stage1-priors}), is derived in
\code{docs/plan/jeffreys\_jeans\_derivation.md}.

\paragraph{Free parameters.}
The default 7-D inference vector
(\code{marginalize\_nuisances=True}, the production setting) is
\begin{equation*}
  \theta = \bigl(\bar V,\;\log_{10} r_s,\;\log_{10}\rho_s,\;\tilde\beta,\;
                 d,\;\varepsilon,\;\vartheta_{1/2}\bigr),
\end{equation*}
where $\vartheta_{1/2}$ is the angular half-light radius (arcmin).
The nuisances $(d,\varepsilon,\vartheta_{1/2})$ enter the likelihood
not only through $r_p$ (eq.~\ref{eq:rp_nuisance}) but also through the
physical projected radius of each star,
\begin{equation}
  R_i = d \cdot \Theta_i \cdot \frac{\pi}{180\times 60}\;[\kpc],
  \label{eq:R-physical}
\end{equation}
where $\Theta_i$ is the angular separation from the galaxy centre.
A larger distance pushes stars to larger physical radii and samples a
different part of the dispersion profile.  The hard constraint
$r_s > r_p$ is enforced per draw inside the likelihood; violators
return $\ln\mathcal{L} = -10^{300}$ and are rejected by the sampler.
The fixed-geometry mode (\code{marginalize\_nuisances=False}) holds
the three nuisances at registry values and reduces $\theta$ to the
4-D vector $(\bar V,\log_{10} r_s,\log_{10}\rho_s,\tilde\beta)$.

\subsection{Priors}
\label{sec:stage1-priors}

All prior bounds and family choices are in
\code{src/dwarfjeans/jeans/priors.py}.

\paragraph{Halo base prior.}
Four base families for $(\ln\rho_s,\ln r_s)$ are registered under
the key \code{prior\_name}:
\begin{description}
  \item[\code{uniform}] linear-uniform on the physical values
    $(r_s,\rho_s)$ with $r_s\in[10^{-2},10]\,\kpc$ and
    $\rho_s\in[10^{4},10^{14}]\,\Msun\,\kpc^{-3}$.  The chain still
    records $(\log_{10} r_s,\log_{10}\rho_s)$.
  \item[\code{loguniform}] uniform on $(\log_{10} r_s,\log_{10}\rho_s)$
    over the same bounds: $\pi(\ln\rho_s,\ln r_s)=\mathrm{const}$.
  \item[\code{jeffreys}] the loguniform base transform plus the
    Jeffreys log-determinant correction $\tfrac{1}{2}\ln D$ added
    to the log-posterior on every likelihood evaluation.
  \item[\code{satgen}] a $\Lambda$CDM-conditioned prior built from
    the SatGen semi-analytic subhalo catalog with a Gaussian conditional
    on $\log_{10}\rho_s\mid\log_{10} r_s$ (Section~\ref{sec:satgen-prior}).
  \item[\code{satgen\_box}] same SatGen catalog but uniform-in-log inside
    the per-bin envelope of populated halos rather than a Gaussian
    conditional (Section~\ref{sec:satgen-box-prior}).
  \item[\code{satgen\_shmr}] per-dwarf SatGen prior obtained by
    reweighting the same halo catalog with precomputed stellar-to-halo-mass-relation
    (SHMR) weights from \code{SatGen\_Dwarf}; same Gaussian-conditional
    functional form as \code{satgen} but with weighted bin moments and
    bin edges (Section~\ref{sec:satgen-shmr-prior}).
\end{description}
The production default is \code{jeffreys}
(\code{scripts/run\_production.py}).

\paragraph{Jeffreys log-determinant.}
The dispersion $\sigma_\mathrm{los}^2(R_i)$ is linear in $\rho_s$ at
fixed $(r_s,\beta)$; its $\ln r_s$-derivative satisfies
$\partial\sigma_\mathrm{los}^2/\partial\ln r_s = \sigma_\mathrm{los}^2 \, T_i$
with auxiliary shape factor $T_i = 3 - \mathcal{Q}(R_i)/\mathcal{P}(R_i)$
(\code{docs/plan/jeffreys\_jeans\_derivation.md}, Steps~2--3).  With per-star weights
$\tilde w_i = A_i^2/(A_i + \sigma_{\varepsilon,i}^2)^2$ and
$A_i = \sigma_\mathrm{los}^2(R_i)$, and membership-weighted sum
$S_0 = \sum_i p_i \tilde w_i$, the $2\times 2$ Fisher matrix in
$(\ln\rho_s,\ln r_s)$ has determinant
\begin{equation}
  D = S_0 \sum_i p_i\,\tilde w_i\,(T_i - \bar T)^2,
  \qquad
  \bar T = \frac{\sum_i p_i\,\tilde w_i\,T_i}{S_0},
  \label{eq:jeffreys-D}
\end{equation}
in the variance form (algebraically equivalent to $S_0 S_2 - S_1^2$
but free of catastrophic cancellation).  The Jeffreys correction is
$\tfrac{1}{2}\ln D$.  When $D \le 0$ or non-finite (all stars at the
same $R_i/r_s$, so $r_s$ is unidentifiable), the correction returns
$-\infty$ and the draw is rejected.  The correction requires $T_i$,
obtained by calling \code{jeans.sigma\_los\_with\_T} instead of
\code{jeans.sigma\_los}; this approximately doubles the cost of each
likelihood evaluation.

\paragraph{SatGen-conditioned prior.}
\label{sec:satgen-prior}
The \code{satgen} family draws $(\log_{10} r_s,\log_{10}\rho_s)$ from
the $\Lambda$CDM-with-tidal-stripping subhalo catalog
\code{m12res8\_10k\_Diemer+scatter\_sim.h5} in the companion
\code{SatGen\_Dwarf} repository ($\sim$$2.4\times 10^6$ surviving and
disrupted subhalos).  Each halo's
$(v_\mathrm{max},r_\mathrm{max})$ is mapped to NFW
$(r_s,\rho_s)$ analytically: the peak of $v_c(r)$ for an NFW profile
occurs at $x_\mathrm{max}\!\equiv\!r_\mathrm{max}/r_s\approx 2.16258$,
the root of $\ln(1+x)=x(1+2x)/[2(1+x)^2]$, so
\begin{align}
  r_s &= r_\mathrm{max}/x_\mathrm{max}
       = 0.46241\,r_\mathrm{max},
       \label{eq:satgen-rs}\\
  \rho_s &= \frac{x_\mathrm{max}^3}{4\pi G\,\mu(x_\mathrm{max})}
            \left(\frac{v_\mathrm{max}}{r_\mathrm{max}}\right)^2
        = 1.72126\,\frac{(v_\mathrm{max}/r_\mathrm{max})^2}{G},
        \label{eq:satgen-rhos}
\end{align}
with $\mu(x)=\ln(1+x)-x/(1+x)$, $v_\mathrm{max}$ converted from
$\kpc\,\mathrm{Gyr}^{-1}$ (SatGen's native units) to $\kms$, and
$G$ in $(\kms)^2\,\kpc/\Msun$.  Halos outside the registry box
$\log_{10}(r_s/\kpc)\in[-2,1]$ are discarded ($\sim$$1.4\times 10^3$
rows of $\sim$$2.4\times 10^6$).

We then build a one-dimensional inverse-CDF table for the SatGen
marginal $p(\log_{10} r_s)$ on a $1024$-node grid, and tabulate
$\mu(\log_{10} r_s)$ and $\sigma(\log_{10} r_s)$ as the per-bin
sample mean and sample standard deviation (\code{ddof=1}) of
$\log_{10}\rho_s$ in $300$ \emph{quantile-cut} bins
(\code{scripts/build\_satgen\_prior\_table.py}; cached at
\code{data/satgen\_prior/m12res8\_diemer\_scatter.npz}).  Bin edges
are placed at equally-spaced quantiles of the empirical
$\log_{10} r_s$ distribution, so every bin contains
$\sim N_\mathrm{total}/N_\mathrm{bins}\approx 8\times 10^3$ halos by
construction; no threshold-based extrapolation or moment-flooring is
needed.  The prior transform draws
\begin{align}
  \log_{10} r_s &\sim F_\mathrm{SatGen}^{-1}(u_1),
        \label{eq:satgen-rs-draw}\\
  \log_{10}\rho_s \mid \log_{10} r_s
   &\sim \mathcal{N}\big(\mu(\log_{10} r_s),\,\sigma(\log_{10} r_s)\big),
        \label{eq:satgen-rhos-draw}
\end{align}
with $\log_{10}\rho_s$ unbounded apart from a $u_2\in[10^{-15},\,1-10^{-15}]$
numerical guard (a $\pm 7.94\sigma$ envelope on the conditional Gaussian
whose excluded mass is $\sim 10^{-15}$).  The conditional tails extend
beyond the registry box on $\rho_s$, but the per-bin scatter is
$\sigma\lesssim 0.5$ so the box edge is reached only on
$\lesssim 10^{-9}$ of draws in the bulk.  An optional
\code{log10\_rs\_min} kwarg re-truncates $F_\mathrm{SatGen}$ to
$[\,\code{log10\_rs\_min},\,1\,]$ and renormalises; it is honoured by
the 4-D builder \code{make\_satgen\_prior\_transform} but \emph{not}
by the 7-D \code{make\_satgen\_prior\_transform\_with\_nuisances}
used in production (hard-coded \code{None}).  $V$ and $\tilde\beta$
priors are unchanged.

The lognormal-in-$\rho_s$ approximation is justified by numerical
diagnostics on the catalog itself, regeneratable via
\code{scripts/build\_satgen\_prior\_table.py}: pooled standardised
residuals across the 2.4M halos have skewness $+0.13$ and excess
kurtosis $+0.21$, with the corresponding QQ relation tracking
$\mathcal{N}(0,1)$ to within $\lesssim 0.1$ in $z$ across $\pm 4\sigma$.
With quantile-cut bins every bin has $\sim 8\times 10^4$ halos, so
per-bin $(\mu,\sigma)$ estimates carry negligible sampling noise and
no bin sits at the unfaithful edge.  Per-bin $\mu$, $\sigma$,
skewness, excess kurtosis, and counts are stored alongside the
lookup arrays in \code{data/satgen\_prior/m12res8\_diemer\_scatter.npz}
(\code{skew\_per\_bin}, \code{kurt\_per\_bin}, \code{n\_per\_bin}); the
top-level \code{metadata} JSON records the pooled moments and the
source-file SHA.  A 2D KDE form is the natural escalation if a future
analysis becomes sensitive to higher-moment structure.

\paragraph{SatGen envelope (box) prior.}
\label{sec:satgen-box-prior}
The \code{satgen\_box} family shares the same SatGen catalog and the
same $(v_\mathrm{max},r_\mathrm{max})\!\to\!(r_s,\rho_s)$ mapping
(eq.~\ref{eq:satgen-rs}, \ref{eq:satgen-rhos}) and the same per-bin
machinery, but replaces the Gaussian conditional in
eq.~\ref{eq:satgen-rhos-draw} with a uniform-in-log envelope.  For
each of the same $300$ quantile-cut $\log_{10} r_s$ bins we record the
empirical extrema
\begin{equation}
  \rho_\mathrm{lo}(k) = \min_{i\in\mathrm{bin}\,k}\log_{10}\rho_{s,i},
  \quad
  \rho_\mathrm{hi}(k) = \max_{i\in\mathrm{bin}\,k}\log_{10}\rho_{s,i}.
\end{equation}
Quantile-cut binning guarantees every bin holds
$\sim 8\times 10^4$ halos, so the empirical $(\rho_\mathrm{lo},\rho_\mathrm{hi})$
are well-defined in every bin and the prior marginal on
$\log_{10} r_s$ is uniform per unit support length across the bins,
\begin{align}
  \log_{10} r_s &\sim \mathrm{Uniform}\!\Big(
        \bigcup_{k:\mathrm{good}} \,[\,\mathrm{edge}_k,\,\mathrm{edge}_{k+1}\,]
        \Big),\\
  \log_{10}\rho_s\mid\log_{10} r_s &\sim
        \mathrm{Uniform}\big(\rho_\mathrm{lo}(k),\,\rho_\mathrm{hi}(k)\big),
\end{align}
where $k$ is the bin containing the drawn $\log_{10} r_s$.  Hard
truncation: no draws land outside the empirical SatGen support, so
no Gaussian-tail leakage past the registry box and no extrapolation
into bins with zero halos.  $V$ and $\tilde\beta$ priors are unchanged.
The new lookup fields \code{bin\_edges\_log10\_rs},
\code{rho\_lo\_log10\_rhos}, \code{rho\_hi\_log10\_rhos}, and the
scalars \code{log10\_rs\_min\_data}, \code{log10\_rs\_max\_data} are
stored in the same cache file alongside the Gaussian-prior fields.
The \code{satgen\_box} family is intended as the closer point of
comparison to importance-reweighted SatGen analyses, which carry
the empirical halo distribution rather than a per-bin Gaussian fit.

\paragraph{SHMR-weighted SatGen prior.}
\label{sec:satgen-shmr-prior}
The \code{satgen\_shmr} family is the per-dwarf refinement of
\code{satgen}: the same Diemer halo catalog
(eq.~\ref{eq:satgen-rs}, \ref{eq:satgen-rhos}) is reweighted by
precomputed importance weights $w_i$ tied to a specific dwarf,
producing one cached lookup table per (SHMR, dwarf) pair instead of
the single galaxy-agnostic table used by \code{satgen}.  The weight
files live in the companion \code{SatGen\_Dwarf} repository under
\code{data/additional/weights\_gc/Diemer/<dwarf>.npz} and are
row-aligned to the halo catalog.  The upstream file contains ten
precomputed SHMRs; the production CLIs
(\code{run\_production.py}, \code{run\_batch.py},
\code{plot\_posteriors.py}) currently wire four of them through the
\code{--shmr} flag: Fattahi et al.\ (2018), Moster et al.\ (2018),
Danieli et al.\ (2023, constant-slope variant), and Kim et al.\
(2024).  The remaining six (Behroozi+13, Moster+13,
Rodriguez-Puebla+17, Behroozi+19, Munshi+21, and the
growing-slope Danieli+23 variant) are available in the build script
but not exposed at the CLI by default.

Each weight $w_i$ is the product of an SHMR likelihood term — a
Bayesian score of the halo's MC-sampled stellar mass against the
dwarf's observed $M_\star$ distribution — and a geometric gate that
zeroes any halo whose galactocentric distance lies outside a factor of
two (in either direction) of the dwarf's observed value.  We consume
the precomputed weight arrays directly; the SHMR formulae,
scatter conventions, and the geometric cut are defined in
\code{SatGen\_Dwarf/python/halo\_weights.py} and not re-derived here.
The geometric gate is dominant: across the 39-dwarf study sample it
zeroes between $28\%$ and $94\%$ of the catalog, leaving
$\sim$$10^5$--$10^6$ non-zero halos per dwarf.  Zero-weight halos
are discarded before any binning, so the weighted moments and
quantile-cut bin edges operate strictly on the geometrically- and
SHMR-consistent sub-population.

The functional form of the prior is identical to \code{satgen}
(eq.~\ref{eq:satgen-rs-draw}--\ref{eq:satgen-rhos-draw}), but every
ingredient is computed with weights.  Bin edges in $\log_{10} r_s$
are placed at uniform quantiles of cumulative weight,
\begin{equation}
  F^{(w)}(x) = \frac{\sum_{i:\,x_i\le x} w_i}{\sum_i w_i},
  \qquad
  \mathrm{edge}_k = (F^{(w)})^{-1}\!\Big(\tfrac{k}{N_\mathrm{bins}}\Big),
\end{equation}
and per-bin moments use the reliability-weighted mean and the
reliability-weight Bessel-corrected variance,
\begin{equation}
  \mu_k = \frac{\sum_{i\in k} w_i\, y_i}{\sum_{i\in k} w_i},
  \quad
  \sigma_k^2 = \frac{\sum_{i\in k} w_i\,(y_i-\mu_k)^2}
                    {\sum_{i\in k} w_i - \sum_{i\in k} w_i^2/\sum_{i\in k} w_i},
  \label{eq:shmr-bessel}
\end{equation}
where $y_i\equiv\log_{10}\rho_{s,i}$.  Eq.~\ref{eq:shmr-bessel}
reduces to \code{np.std(..., ddof=1)} when all weights are equal, and
inflates the denominator by the effective sample size for unequal
weights, matching the \code{satgen} convention.

We compute the global effective sample size
$\mathrm{ESS} = (\sum_i w_i)^2/\sum_i w_i^2$ for each dwarf and
default to $N_\mathrm{bins}=300$; if $\mathrm{ESS}/N_\mathrm{bins}<30$
we adaptively reduce $N_\mathrm{bins}$ to
$\max(50,\lfloor\mathrm{ESS}/30\rfloor)$ so each bin retains at least
$\sim 30$ effective halos.  In the Fattahi+18 build of the
39-dwarf sample only Hydrus~I triggered the fallback
($N_\mathrm{bins}=292$); ESS ranged from $\sim$$9\times 10^3$ to
$\sim$$1.3\times 10^5$.  Tables and per-dwarf metadata
(\code{n\_halos\_nonzero}, \code{zero\_weight\_fraction},
\code{ess}, \code{n\_bins}) are cached at
\code{data/satgen\_prior/<shmr>/<lvdb\_key>.npz}; the builder is
\code{scripts/build\_satgen\_shmr\_prior\_tables.py} and the loader
shares the \code{satgen} code path
(\code{make\_satgen\_shmr\_prior\_transform} in
\code{src/dwarfjeans/jeans/priors.py} is a thin wrapper that
resolves the per-dwarf table path and delegates to
\code{make\_satgen\_prior\_transform}).  The
\code{satgen\_shmr} prior has no Walker constant-$\sigma$ analog on
$\sigma_\mathrm{los}$, so the companion in
\code{run\_production.py} falls back to \code{jeffreys} for that
inference, as for \code{satgen}/\code{satgen\_box}.

\paragraph{Anisotropy and systemic priors.}
$\tilde\beta$ is flat on $[-0.95,\,1)$ (eq.~\ref{eq:beta-tilde}); the
open upper bound avoids the radial-orbit singularity at $\beta=1$.
$\bar V$ is flat on $[V_\mathrm{centre} - 10,\,V_\mathrm{centre} + 10]\,\kms$,
where $V_\mathrm{centre}$ is the inverse-variance-weighted mean of the
post-selection velocities,
$V_\mathrm{centre} = \sum_i V_i/\sigma_{\varepsilon,i}^2 \big/ \sum_i 1/\sigma_{\varepsilon,i}^2$.
Centering on the data rather than on the registry
\code{vlos\_systemic\_kms} avoids stale/sample-mismatch failures
(Pegasus III hit the lower wall of a registry-centered prior in QA).

\paragraph{Nuisance priors (7-D mode).}
When \code{marginalize\_nuisances=True}, three additional parameters
are sampled with independent Gaussian priors derived directly from
the registry.  The interim symmetric-Gaussian form below matches the
current implementation; the design target (split-normal priors for
asymmetric LVDB error fields) and the rationale for both choices are
in \code{docs/plan/uncertainty\_conventions.md}.
\begin{align}
  d &\sim \mathcal{N}(\mu_d,\,\sigma_d), \label{eq:prior-d}\\
  \varepsilon &\sim \mathcal{N}(\mu_\varepsilon,\,\sigma_\varepsilon)
              \text{ truncated to } [0,1), \label{eq:prior-eps}\\
  \vartheta_{1/2} &\sim \mathcal{N}(\mu_{\vartheta},\,\sigma_{\vartheta}),
              \label{eq:prior-rhalf}
\end{align}
with $\sigma_d = (\ln 10/5)\,d\,\sigma_\mu$ from the registry
distance modulus and its uncertainty
(\code{run\_production.\_registry\_nuisance\_priors}).  For
ellipticity-missing galaxies, $\sigma_\varepsilon$ is set to half
the registry upper limit.

\subsection{Sampler}
\label{sec:stage1-sampler}

\paragraph{Algorithm.}
Static nested sampling with \textsc{dynesty} (Speagle 2020); bounding
method \code{multi}, proposal method \code{rwalk}.  Production
defaults: $N_\mathrm{live}=500$ and stopping criterion
$\Delta\ln\mathcal{Z}<0.1$.  The random state is seeded via
\code{np.random.default\_rng(rseed)} with default \code{rseed=0}.

\paragraph{Post-processing.}
After \code{sampler.run\_nested}, normalised log-weights are
$\log w_k = \log\mathrm{wt}_k - \ln\mathcal{Z}$, and
\code{dynesty.utils.resample\_equal} produces an equal-weight chain
\code{samples\_eq} from which all downstream quantities are derived.
The driver returns the raw and equal-weight samples, log-weights,
$\ln\mathcal{Z}$ and its uncertainty, and the parameter-name tuple.

\paragraph{Derived chains.}
The production driver
\code{scripts/run\_production.py} pushes \code{samples\_eq} through:
per-sample $r_p$ and the radii
$R_{1/2,\mathrm{2D}} = r_p$ and $r_{1/2,\mathrm{3D}} = 1.30477\,r_p$;
NFW enclosed masses $M(R_{1/2,\mathrm{2D}})$ and
$M(r_{1/2,\mathrm{3D}})$ via \code{jeans.nfw\_M};
$\sigma_\mathrm{los}$ at $R_{1/2,\mathrm{2D}}$ (thinned to 2000 draws);
and $J/D$ chains at four angles each (thinned to 500 draws,
Section~\ref{sec:stage3:chain}).  Outputs land in
\code{results/production/<lvdb\_key>/<prior>/} (overwritten by each
run) as
\code{posterior\_samples.npz}, a per-quantity $16/50/84$-percentile
\code{summary.csv}, and an \code{audit.json} recording registry
inputs, selection policy, and dynesty configuration.

\subsection{Walker constant-$\sigma$ baseline}
\label{sec:stage1:walker}

Alongside the Jeans posterior, the production driver also evaluates
a radius-independent Walker (2006) dispersion as a model-free
cross-check.  This is the canonical comparison number quoted across
the dwarf-galaxy literature (P\&S 2018, Walker 2006), so reporting it
in the same \code{summary.csv} as the Jeans result makes paper-vs-paper
checks one-line.

\paragraph{Likelihood.}
A single global $\sigma$ replaces the radial profile
$\sigma_\mathrm{los}(R)$.  The per-star membership-weighted Gaussian
matches Eq.~\eqref{eq:lnlike} stripped of the $R$-dependence
($s_i^{2}\to\sigma^{2}+\varepsilon_i^{2}$ instead of
Eq.~\eqref{eq:stot}):
\begin{equation}
\ln L_i \;=\; p_i\,\Big[\,
  -\tfrac{1}{2}\ln\!\big[2\pi(\sigma^{2} + \varepsilon_i^{2})\big]
  -\tfrac{1}{2}\,\frac{(V_i - V_\mathrm{sys})^{2}}{\sigma^{2} + \varepsilon_i^{2}}
\,\Big].
\end{equation}
Inference is on the two parameters $(V_\mathrm{sys}, \sigma)$ only;
no halo parameters and no nuisances enter.

\paragraph{Priors.}
Uniform on $V_\mathrm{sys}\in[V_\mathrm{centre}\pm 10\;\mathrm{km\,s^{-1}}]$
($V_\mathrm{centre}$ is the IVW mean of the post-selection velocities,
matching the Stage~2 prior; see the anisotropy-and-systemic-priors
paragraph).  On $\sigma$ we use the
\emph{proper} (Fisher-determinant) Jeffreys prior, computed at each
grid point from the diagonal Fisher information $I_{\bar V\bar V} =
1/\sigma_i^{2}$ and $I_{\sigma\sigma} = 2\sigma^{2}/\sigma_i^{4}$
with $\sigma_i^{2} = \sigma^{2} + \varepsilon_i^{2}$:
\begin{equation}
p_J(\sigma) \;\propto\; \sigma\,
\sqrt{\Big[\sum_i p_i/\sigma_i^{2}\Big]\Big[\sum_i p_i/\sigma_i^{4}\Big]},
\end{equation}
which is $\bar V$-independent and truncated to
$\log_{10}\sigma\in[-2, 2]$ (i.e.\ $\sigma\in[0.01, 100]\;\mathrm{km\,s^{-1}}$).
The proper Jeffreys form is preferred over the bare log-uniform
because the latter piles posterior mass at $\sigma\!\ll\!\varepsilon_i$
where the likelihood is $\sigma$-independent.

\paragraph{Implementation and reporting.}
The 2D posterior on $(V_\mathrm{sys}, \sigma)$ is evaluated on a
$400\times400$ grid in \code{constant\_sigma\_inference}.  Marginals
on $V_\mathrm{sys}$ and $\sigma$ are integrated by trapezoid; the
$16/50/84$ percentiles are read off the cumulative.  The driver
records the $\sigma$ marginal as \code{sigma\_los\_walker\_kms} in
\code{summary.csv}.  Runtime is sub-second per galaxy and is
independent of the Jeans inference.

% ===================================================================
\section{Stage 2: Population Monte Carlo calibration}
\label{sec:stage2}

The MC test design (truth grids, realisation count, bias / coverage
metrics, rerun procedure) is documented in
\code{docs/plan/README\_mc\_test.md}; the calibration role within the
pipeline is in \code{docs/plan/stage2.md} (``MC recovery test'' section).

\subsection{Motivation}

Before applying the Jeans inference to real kinematic data, the full
chain --- mock generation, Jeans projection, likelihood, sampler,
posterior summarisation --- is validated end-to-end on synthetic
catalogs at known truth.  Unit tests verify the mathematics of
individual components; the population MC verifies that the complete
chain is unbiased and well-calibrated on realistic inputs.

\subsection{Mock galaxy generator}
\label{sec:stage2:mock}

The mock generator is implemented in
\code{src/dwarfjeans/mocks/galaxy.py}.  Each realisation produces a
per-star catalog $(R_i, V_i, \sigma_{\varepsilon,i}, p_i)$ matching
the schema of the real kinematic data; the truth model is an NFW halo
$(r_s,\rho_s)$ with a Plummer tracer of scale $r_p$ and constant
anisotropy $\beta$.

\paragraph{Projected radii.}
Stars are drawn from the 2D Plummer surface density
$\Sigma(R)\propto(1+R^2/r_p^2)^{-2}$ via inverse-CDF sampling; the
CDF $F(<R) = R^2/(R^2+r_p^2)$ inverts analytically:
\begin{equation}
  R_i = r_p\sqrt{\frac{u_i}{1-u_i}},\qquad u_i\sim\mathrm{Uniform}(0,1).
\end{equation}
Draws with $R_i > 5\,r_p$ are rejected and redrawn, matching the
membership-cut footprint of real spectroscopic surveys.

\paragraph{Line-of-sight velocities.}
For each $R_i$, $\sigma_\mathrm{los}(R_i)$ is evaluated at truth and
\begin{equation}
  V_i \sim \mathcal{N}\!\bigl(V_\mathrm{sys},\,
    \sigma_\mathrm{los}^2(R_i)+\sigma_{\varepsilon,i}^2\bigr).
\end{equation}
Membership probabilities are set to $p_i = 1$ (post-membership-cut).

\subsection{Asimov realisation}
\label{sec:stage2:asimov}

A single Asimov realisation
(\code{src/dwarfjeans/mocks/galaxy.make\_asimov\_galaxy}; driver
\code{tests/integration/run\_ufd\_population.py --asimov}) isolates
prior and projection bias from sampling noise.  Star positions are
placed at the equal-probability stratified midpoints of the
truncated Plummer CDF on $R\in[0,5\,r_p]$:
\begin{equation}
  u_i = \frac{i - \tfrac12}{N},\quad
  R_i = r_p\sqrt{\frac{u_i\,F_\mathrm{max}}{1 - u_i\,F_\mathrm{max}}},
  \qquad i=1,\ldots,N,
\end{equation}
where $F_\mathrm{max} = (5r_p)^2/[(5r_p)^2 + r_p^2]$.  The Asimov
likelihood replaces the squared velocity residual by its
expectation $\sigma_\mathrm{tot,truth}^2(R_i)$ under the truth model,
so the MLE sits exactly at truth.  $V_\mathrm{sys}$ does not enter
the Asimov likelihood and its posterior equals the prior; it is
flagged \code{prior\_only} and excluded from $z$-score diagnostics.

\subsection{Population MC test}
\label{sec:stage2:mc}

Implemented in \code{tests/integration/run\_ufd\_population.py}.
Fifteen independent realisations are generated at the same truth:
\begin{center}
\begin{tabular}{ll}
$r_s = 0.3\,\kpc$ & $\rho_s = 3\times 10^{8}\,\Msun\,\kpc^{-3}$ \\
$r_p = 0.05\,\kpc$ & $\beta = 0$ \\
$N_\star = 30$ & $\sigma_\varepsilon = 2\,\kms$
\end{tabular}
\end{center}
Each realisation uses a distinct random seed.  Inference uses
\code{nlive}=300, \code{dlogz}=0.5, log-uniform prior, and uniform
sampling (\code{sample="unif"}) for speed --- distinct from the
production driver, which uses \code{sample="rwalk"} with the same
\code{multi}-ellipsoidal bounds.  Calibration
diagnostics are computed for each parameter
$\theta\in\{\bar V,\,\log_{10} r_s,\,\log_{10}\rho_s,\,\tilde\beta,\,
\log_{10}(\rho_s r_s^3)\}$:
\begin{itemize}
  \item asymmetric $z$-score
        $z = (\mathrm{median}-\theta_\mathrm{truth})/\sigma_\pm$,
  \item 68\% and 95\% credible-interval coverage,
  \item Kolmogorov--Smirnov $p$-value of $\{z\}$ against
        $\mathcal{N}(0,1)$,
  \item mean $z$, std$(z)$, and median bias.
\end{itemize}

\paragraph{Calibration results (Jeffreys prior, current default).}
The 15-realisation MC was rerun with \code{prior\_name="jeffreys"} to
match the production default (\code{run\_ufd\_population.py}). Per-parameter
diagnostics:

\begin{center}
\begin{tabular}{lrrrrrr}
\hline
Parameter & mean$(z)$ & std$(z)$ & med.\ bias & cov$_{68\%}$ & cov$_{95\%}$ & KS $p$ \\
\hline
$\bar V$                  & $-0.08$ & $0.93$ & $-0.08$ & $87\%$ & $93\%$  & $0.75$ \\
$\log_{10} r_s$           & $+0.72$ & $0.60$ & $-0.65$ & $60\%$ & $100\%$ & $0.0002$ \\
$\log_{10}\rho_s$         & $-0.77$ & $0.63$ & $+1.10$ & $60\%$ & $100\%$ & $0.0003$ \\
$\tilde\beta$             & $+0.17$ & $0.81$ & $+0.00$ & $87\%$ & $100\%$ & $0.73$ \\
$\log_{10}(\rho_s r_s^3)$ & $+0.65$ & $0.56$ & $-0.86$ & $67\%$ & $100\%$ & $0.0002$ \\
$\log_{10} M(r_{1/2,2\mathrm{D}})$ & $-0.83$ & $0.82$ & $+0.13$ & $53\%$ & $100\%$ & $0.013$ \\
$\log_{10} M(r_{1/2,3\mathrm{D}})$ & $-0.61$ & $0.79$ & $+0.07$ & $60\%$ & $100\%$ & $0.071$ \\
\hline
\end{tabular}
\end{center}

The well-constrained derived quantity remains $M(r_{1/2,\mathrm{3D}})$
(typical $1\sigma$ width $\approx 0.12$~dex, smaller than the
loguniform value of $0.15$~dex), but under Jeffreys it picks up a
$\sim 0.07$~dex high-side median bias and a tighter-than-nominal
$z$-distribution ($\mathrm{std}(z) = 0.79$, $\mathrm{cov}_{68\%}=60\%$,
KS $p=0.07$). The individual halo parameters $\log_{10} r_s$,
$\log_{10}\rho_s$, $\log_{10}(\rho_s r_s^3)$ are now meaningfully
biased ($|\mathrm{med.\ bias}| = 0.6$--$1.1$~dex) with strongly
non-normal $z$-distributions (KS $p < 10^{-3}$).

This is a marked departure from the historical loguniform-prior
result, where $M(r_{1/2,\mathrm{3D}})$ was recovered with median bias
$<0.01$~dex, $\mathrm{std}(z)\approx 1.1$, and KS $p\approx 0.7$. The
Jeffreys term shifts the joint posterior peak toward higher
$\rho_s$ / lower $r_s$ in the small-$x$ NFW regime relevant for
UFDs ($r_{1/2}/r_s\approx 0.2$), where the data only constrain the
enclosed-mass combination $M \propto \rho_s r_s r^2$. The inflated
peak shifts $M(r_{1/2})$ slightly high and tightens the marginal
posterior beyond the realisation-to-realisation spread, producing
the observed under-dispersion.

\paragraph{Prior note.}
The 15-realisation MC is now run under the production prior
(\code{prior\_name="jeffreys"}) to keep the calibration aligned with
production. The historical log-flat numbers (median bias $<0.01$~dex,
$\mathrm{std}(z)\approx 1.1$, KS $p\approx 0.7$) remain accessible by
overriding to \code{prior\_name="loguniform"}.

% ===================================================================
\section{Stage 3: $J$- and $D$-factor extraction}
\label{sec:stage3}

The canonical extended treatment of the $J/D$-factor pipeline (the
chain-push design, tidal-radius source choice, MC recovery plan, and
deviations from P\&S 2018) lives in \code{docs/plan/stage3.md}; this
section summarizes the as-implemented integrals and reporting angles.

\subsection{Integral definitions}
\label{sec:stage3:integrals}

The $J$- and $D$-factors quantify the line-of-sight integral of
$\rho^2$ and $\rho$, respectively, over a solid angle
$\Delta\Omega(\theta_\mathrm{max})$ centred on the target
(\code{src/dwarfjeans/jd/factors.py}):
\begin{align}
  J(\theta_\mathrm{max}) &= \int_{\Delta\Omega}\int_\mathrm{l.o.s.}
    \rho^2\!\bigl(r(\ell,\Omega)\bigr)\,d\ell\,d\Omega, \\
  D(\theta_\mathrm{max}) &= \int_{\Delta\Omega}\int_\mathrm{l.o.s.}
    \rho\!\bigl(r(\ell,\Omega)\bigr)\,d\ell\,d\Omega.
\end{align}
In the small-angle approximation (valid for $\theta_\mathrm{max}\lesssim 1^\circ$),
$R = d\,\theta$ and $d\Omega \approx R\,dR/d^2$, giving
\begin{align}
  J(\theta_\mathrm{max}) &= \frac{2\pi}{d^2}
    \int_0^{R_\mathrm{max}} R\,I_2(R;\,r_t)\,dR, \\
  D(\theta_\mathrm{max}) &= \frac{2\pi}{d^2}
    \int_0^{R_\mathrm{max}} R\,I_1(R;\,r_t)\,dR,
\end{align}
where $R_\mathrm{max} = \min(d\,\theta_\mathrm{max},\,r_t)$ and the
line-of-sight column at impact parameter $R$ for an NFW halo with
hard tidal truncation at $r_t$ is
\begin{equation}
  I_n(R;\,r_t) = 2\int_R^{r_t}
    \frac{\rho^n(r)\,r}{\sqrt{r^2-R^2}}\,dr,\qquad n=1,2.
\end{equation}
The integrable $1/\sqrt{r^2-R^2}$ singularity is removed by
$r^2 = R^2 + u^2$, giving a smooth integrand on
$u\in[0,u_\mathrm{max}(R)]$ with $u_\mathrm{max}(R)=\sqrt{r_t^2-R^2}$.
Internally the pipeline works in $(\Msun,\kpc)$ and converts to
P\&S~2018 reporting units ($\mathrm{GeV}^2/\mathrm{cm}^5$ for $J$,
$\mathrm{GeV}/\mathrm{cm}^2$ for $D$) at output.

\subsection{Numerical implementation}
\label{sec:stage3:numerics}

The $u$-grid is log-spaced from $10^{-6}\,r_s$ to $u_\mathrm{max}(R)$
with an explicit $u=0$ endpoint prepended (where the integrand is
largest); trapezoidal quadrature is applied.  The outer $R$-grid is
log-spaced from $10^{-6}\,r_s$ to $R_\mathrm{max}$ with the on-axis
point $R=0$ excluded from the log grid (where $I_n(0)$ hits
$\rho(0)=\infty$ for an NFW); the integrand $R\cdot I_n(R)$ is zero
at $R=0$, so an explicit $(R=0,\mathrm{integrand}=0)$ endpoint is
prepended.  Production defaults from \code{J\_D\_factors} are $n_R=128$,
$n_u=512$, calibrated against \code{scipy.integrate.quad} to better
than $10^{-3}$ relative error over the MW-dwarf parameter range.
\code{summarize\_jd} (used by population-MC summarisation) defaults
to $n_R=64$, $n_u=128$ and thins the chain to 500 samples;
\code{analyze\_asimov.py} runs at $n_R=96$, $n_u=256$.

\subsection{Tidal angle and reporting angles}
\label{sec:stage3:angles}

The characteristic integration angle for $J$ follows the
Wolf~et~al.\ (2010) critical-angle convention:
\begin{equation}
  \alpha_c = \frac{2\,r_{1/2,\mathrm{3D}}}{d},
\end{equation}
computed via \code{alpha\_c\_radians} in \code{factors.py}.  Each
chain is evaluated at four angles:
\begin{align}
  J&:\quad \theta_\mathrm{max} \in \{0.1^\circ,\,0.2^\circ,\,0.5^\circ,\,\alpha_c\},\\
  D&:\quad \theta_\mathrm{max} \in \{0.1^\circ,\,0.2^\circ,\,0.5^\circ,\,\alpha_c/2\}.
\end{align}
The $\alpha_c/2$ choice for the $D$ natural angle reflects the
shallower projected $\rho$ profile relative to $\rho^2$.  In
nuisance-marginalised mode $\alpha_c$ becomes a per-draw quantity;
the fixed angles vary only through the sampled $d$.

\subsection{Chain push}
\label{sec:stage3:chain}

The Stage~1 equal-weight chain
$(\bar V^{(k)},\,\log_{10} r_s^{(k)},\,\log_{10}\rho_s^{(k)},\,\tilde\beta^{(k)})$
is processed sample-by-sample in \code{scripts/run\_production.py}
via per-draw, per-angle calls to \code{J\_D\_factors}
(\code{factors.py}).  Per draw: $(r_s,\rho_s)$ are extracted; $d$
comes from the nuisance chain (production) or is held at truth (MC
validation); the tidal radius $r_t$ is hard-coded to $1\,\kpc$ in
the production driver, since the input chain carries no host-orbit
information from which to derive a Springel~et~al.\ 2008 tidal
radius. \code{J\_D\_factors} returns $(J^{(k)}, D^{(k)})$ in internal
units.  The chains are summarised per galaxy as median +
$q_{16}/q_{84}$ + 1D-KDE MAP.

\subsection{MC recovery}
\label{sec:stage3:mc}

The 15 mocks of Section~\ref{sec:stage2:mc} are pushed through $J/D$
with $d=30\,\kpc$ and $r_t=1\,\kpc$ held at truth (synthetic data
carry no host-orbit information for a Springel~et~al.\ 2008 $r_t$).

\emph{Population-level J/D diagnostics are pending under the
Jeffreys-prior MC.} The historical numbers below were produced by the
former \code{run\_jd\_summary.py} tool under the loguniform prior; the
current \code{run\_ufd\_population.py} computes the halo recovery only.
Reinstating per-realisation J/D summaries (via a \code{summarize\_jd}
call inside the MC loop, writing \code{ufd\_pop\_jd\_diagnostics.json})
is a follow-up. Historical loguniform numbers, retained as a baseline:
\textbf{$D$-factor recovery was clean} at all four angles (median
bias $\le 0.03$~dex, std$(z)\le 1.15$, $\max|z|\approx 2.1$, KS
$p\in[0.37,0.79]$, mean $1\sigma$ posterior widths $0.15$--$0.36$~dex);
\textbf{$J$-factor recovery was unbiased to $\sim 0.1$~dex} with
std$(z)\le 1.18$, $\max|z|\approx 2.3$, and a small high-side median
bias of $+0.08$ to $+0.12$~dex. The Asimov decomposition below is
re-verified under Jeffreys and matches the historical loguniform
numbers to within $\sim 0.01$~dex.

\subsection{Asimov-bias decomposition}
\label{sec:stage3:asimov}

The Asimov chain (1803 equal-weight samples;
\code{tests/integration/analyze\_asimov.py}) isolates the origin of
the $J$-factor median offset.

\paragraph{(a) MAP vs.\ median vs.\ truth.}
Pushing through $J(\alpha_c)$:
\begin{center}
\begin{tabular}{lc}
\hline
Estimator & Offset from truth (dex) \\
\hline
1D-KDE MAP of $\log_{10} J(\alpha_c)$ & $+0.037$ \\
Chain median of $\log_{10} J(\alpha_c)$ & $+0.162$ \\
Chain mean of $\log_{10} J(\alpha_c)$ & $+0.214$ \\
\hline
\end{tabular}
\end{center}
The posterior peak (MAP) sits within $0.02$--$0.13$~dex of truth at
all four angles; the offset lives entirely in the median of a
positively skewed transformation distribution.

\paragraph{(b) Small-$x$ analytic check.}
In the small-$x$ NFW limit, $J\propto\rho_s^2 r_s^3$.  The chain
median of $2\log_{10}\rho_s + 3\log_{10} r_s$ is $+0.42$~dex,
against an actual $J(0.1^\circ)$ bias of $+0.13$~dex.  Per-sample
regression of $\Delta\log_{10} J$ on the small-$x$ prediction gives
slopes $0.29$--$0.53$ (not unity) and $r^2 = 0.08$--$0.52$ across
the four angles --- the log-linear small-$x$ attribution does not
hold over the full chain support.

\paragraph{(c) 1D-projection decomposition at $\alpha_c$.}
\begin{center}
\begin{tabular}{lc}
\hline
Projection & Median bias (dex) \\
\hline
$\log r_s$ only ($\log\rho_s = \mathrm{truth}$) & $+0.45$ \\
$\log\rho_s$ only ($\log r_s = \mathrm{truth}$) & $-0.32$ \\
Sum & $+0.13$ \\
Full joint chain & $+0.16$ \\
Non-additive remainder & $+0.03$ \\
\hline
\end{tabular}
\end{center}
The full bias is a small residual between two large opposite-sign
1D-projection offsets.  The non-additivity is negligible
($0.02/0.16$~dex), indicating that the chain ridge is approximately
log-linearly aligned with the $J$ surface.  The $J/D$ bias ratio
across angles is $2.1$--$2.8$, consistent with $\rho^2$ vs.\ linear
$\rho$ kernels both experiencing the same posterior-curvature skew.
The pipeline publishes median + $q_{16}/q_{84}$ + 1D-KDE MAP for
every $\log_{10} J(\theta)$ and $\log_{10} D(\theta)$, so the
median--MAP gap is visible to downstream consumers without any
correction applied.

% ===================================================================
\section{Tests}
\label{sec:tests}

\subsection{Unit tests}
\label{sec:unit}

Unit tests reside in \code{tests/unit/} and are run with
\textsc{pytest}.  Each file targets one module or contract.

\paragraph{\code{tests/unit/test\_jeans.py}} verifies the Jeans
solver across four categories: (1) the NFW auxiliary $g(x)$ via the
small-$x$ Taylor series and the stable-regime $\ln(1+x)$ form;
(2) a $\tilde\beta$ prior-edge smoke test producing finite, positive
$\sigma_\mathrm{los}$ with zero \code{RuntimeWarning}s across
$\tilde\beta\in\{-0.95,-0.5,0,+0.5,+0.95\}$, including the
tangential extreme $\beta\approx -38$; (3) physical sanity ---
doubling $\rho_s$ multiplies $\sigma_\mathrm{los}$ by $\sqrt 2$ to
$10^{-3}$ and $\sigma_\mathrm{los}(r_p)$ falls within a factor of
three of the Wolf~et~al.\ (2010) estimator; (4) a 100-star
likelihood call completes in under 50~ms on a single core.

\paragraph{\code{tests/unit/test\_jeans\_vs\_quad.py}} cross-checks
the production log-grid integrator against \code{scipy.integrate.quad}
(treated as ground truth, $\sim 10^4\times$ slower).  Inner Jeans
integrands and line-of-sight projections are compared across eight
parameter combinations spanning the UFD and classical regimes at
isotropic, radial, and tangential $\beta$.  Tolerances $10^{-3}$
(inner) / $10^{-2}$ (projection).  Run on demand before merging
changes to integration tolerances or grid sizes.

\paragraph{\code{tests/unit/test\_multi\_epoch.py}} verifies the
multi-epoch primitives.  IVW combination is checked against the
textbook formula and against a closed-form expectation for
$\sigma_\mathrm{sys}>0$ added in quadrature post-combine.  The
$\chi^2$ test is checked at zero scatter (no flag), at a $5\sigma$
outlier (flag), and for single-epoch input ($\chi^2,p=\mathrm{NaN}$,
no flag).  The strict-deconvolution path is verified to equal the
as-statistical path when $\sigma_\mathrm{sys}=0$, to return a
strictly larger $\sigma_{\bar v}$ at fixed published total error,
to raise \code{ValueError} when the implied $\sigma_\mathrm{stat}$
would be imaginary, and (single-epoch) to return
$\sigma_{\bar v}=\sigma_\mathrm{total}$ exactly.  Strict $\chi^2$
uses $\sigma_\mathrm{stat}$ rather than $\sigma_\mathrm{total}$,
producing a larger statistic.

\paragraph{\code{tests/unit/test\_default\_combiner.py}} verifies the
zero-point offset hook.  Empty offsets reproduce the no-shift IVW.
With offsets, velocities are shifted additively before IVW and the
output recovers the common-zero-point $\bar v$.  The combiner raises
\code{KeyError} on missing \code{Inst} column, raises on unknown
instrument tag, and does not mutate the caller's input array.

\paragraph{\code{tests/unit/test\_combiner\_dispatch.py}} checks the
per-paper combiner registry.  Each of the six per-epoch bibcodes
(Li+2017/2018, Chiti+2022/2023, Simon+2020, Hansen+2024) dispatches
to a dedicated handler.  An unknown bibcode falls back to the
default combiner with $\sigma_\mathrm{sys}=0$ and
$p_\mathrm{threshold}=0.01$.

\paragraph{\code{tests/unit/test\_preprocess.py}} exercises
\code{prepare\_jeans\_input}.  Per-star catalogs bypass the
combiner; per-epoch catalogs invoke the registered handler followed
by selection.  A two-star, two-epoch example verifies that a
variable star is dropped by the variability cut.  Policies are
threaded through and appear in the audit dict.  Missing
\code{source\_paper\_bibcode} on a per-epoch catalog and unknown
\code{catalog\_granularity} both raise \code{ValueError}.  A
parametrised smoke test runs every per-epoch catalog through the
full pipeline.

\paragraph{\code{tests/unit/test\_priors.py}} checks the prior
registry.  Exactly the three keys \code{uniform},
\code{loguniform}, \code{jeffreys} are registered; an unknown key
raises.  The log-uniform transform is checked byte-for-byte over
1000 unit-cube draws.  The Jeffreys log-correction returns
$-\infty$ when all $T_i$ are equal (degenerate Fisher determinant)
and a finite value otherwise; the \code{needs\_T} flag is checked,
and the correction returns exactly zero for the uniform and
log-uniform priors.

\paragraph{\code{tests/unit/test\_selection.py}} verifies
\code{select\_jeans\_stars}.  Membership, radial, and variability
cuts are exercised in isolation and in combination, including a
regression for a bug where uniform $p\in\{0,1\}$ membership was
mistaken for a pre-applied cut.  Disabling the variability cut
retains all stars.  Missing \code{R}, invalid $r_{1/2}$, and
per-epoch input each raise.  An \code{NpzFile} compatibility test
verifies that \code{np.load} output is accepted directly.

\subsection{Integration tests}
\label{sec:integration}

\subsubsection{Segue~1 end-to-end test}
\label{sec:segue1}

\code{tests/integration/run\_segue1.py} runs the complete pipeline on
Segue~1 and compares the output to published results.  Full toggles,
results history, and the P\&S~2018 comparison table live in
\code{docs/plan/segue1\_test.md}; this section summarizes the test
spec.  Three per-star
data sources are toggled by the \code{SOURCE} variable: the
Simon~et~al.\ (2011) VizieR catalog with $B_\mathrm{pr}>0.8$
(\code{simon}); the Pace pre-combined Bayes file with $p\ge 0.8$
(\code{pace}); and the Geha~et~al.\ (2026) catalog staged through
the production \code{prepare\_jeans\_input} pipeline with
$p_\mathrm{mem,novar}=1$ and the
$R<2\,r_{1/2,\mathrm{maj}}\,\sqrt{1-\varepsilon}\,(4/3)$
sphericalised-3D-Plummer cut (\code{geha}); the variability removal is
already folded into $p_\mathrm{mem,novar}$ upstream in Paper~I.  For \code{simon} / \code{pace},
nuisance priors follow P\&S~2018: $d=23\pm 2\,\kpc$,
$r_{1/2,\mathrm{2D}}=21\pm 5\,\mathrm{pc}$
($r_h=4.31\pm 1.03\,\mathrm{arcmin}$ with $\varepsilon=0.47\pm 0.11$).
For \code{geha}, nuisance priors are taken from LVDB~v1.0.5.  The
reference is $\sigma_\mathrm{los}=3.1^{+0.9}_{-0.8}\,\kms$ from
P\&S~2018 Table~A1.  All three paths run 7-D dynesty inference and
produce posterior samples, a summary CSV, and diagnostic plots in
\code{results/tests/segue1/}.

\subsubsection{UFD population MC recovery}
\label{sec:ufd_population}

\code{tests/integration/run\_ufd\_population.py} performs the
15-realisation MC of Section~\ref{sec:stage2:mc}.  All realisations
share the same injected truth; realisations differ only in the random
seed for stellar positions and velocity scatter.  Each realisation
runs dynesty (\code{nlive}=300, \code{dlogz}=0.5, log-uniform prior)
or reloads a cached chain from
\code{results/tests/ufd\_population/}.  Population diagnostics for
five parameters
($\bar V$, $\log_{10} r_s$, $\log_{10}\rho_s$, $\tilde\beta$,
$\log_{10}(\rho_s r_s^3)$) are written to
\code{results/tests/ufd\_population/ufd\_pop\_diagnostics.json}.
A single Asimov realisation is run via the \code{--asimov} flag.

\subsubsection{Asimov bias decomposition}
\label{sec:asimov}

\code{tests/integration/analyze\_asimov.py} loads the chain produced
by \code{--asimov} and performs six diagnostics: (1)
$M(r_{1/2,\mathrm{3D}})$ recovery (median, mean, 1D-KDE peak vs.\
truth, plus skewness and excess kurtosis); (2) full-chain push
through $J$ and $D$ at four angles each; (3) small-$x$ analytic
test on the $2\log_{10}\rho_s + 3\log_{10} r_s$ proxy; (4)
per-sample $\Delta\log_{10} J$ regression on the small-$x$ prediction
(slope and $R^2$); (5) marginal-projection decomposition at
$\alpha_c$ holding one channel at truth and pushing the other; (6)
prior-edge fraction within $\{0.05,0.10,0.20\}$ of the
$\log_{10} r_s$ prior boundaries.  Outputs are saved to
\code{results/tests/ufd\_population/asimov\_chain\_diagnostics.npz};
the headline numbers in Section~\ref{sec:stage3:asimov} come from
this run.

% ===================================================================
\section{Stage 2 systematic corrections}
\label{sec:stage2-corrections}

Two corrections to the Stage~2 per-star likelihood live in this
section.  The Kaplinghat--Strigari perspective-motion correction
(\S\ref{sec:open:perspective}) is implemented end-to-end, with PM
uncertainty marginalised inside Stage~2.  Velocity gradients
(\S\ref{sec:open:gradients}) remain an open issue.

\subsection{Perspective-motion correction}
\label{sec:open:perspective}

For satellites with measured bulk proper motion the Kaplinghat--Strigari
perspective correction (Walker et al.\ 2008, Appendix; Kaplinghat \&
Strigari 2008) adjusts each per-star $v_\mathrm{los}$ by the projection
of the galaxy's plane-of-sky bulk velocity onto the star's line of
sight, which differs from the line of sight to the galaxy centre for
stars at non-zero angular separation.  In the small-angle limit
appropriate to our sample (fields of $\lesssim 1^\circ$ across),
\begin{equation}
\label{eq:persp}
\Delta v_\mathrm{persp}(\alpha,\delta) = A\,d\,\bigl(\mu_{\alpha\ast}\,\Delta\alpha\cos\delta_0 + \mu_\delta\,\Delta\delta\bigr),
\end{equation}
with $A = 4.74047~\mathrm{km\,s^{-1}}\,(\mathrm{mas\,yr^{-1}\,kpc})^{-1}$,
$d$ in kpc, $\mu_{\alpha\ast} = \mu_\alpha\cos\delta$ in the Gaia
convention, and $(\Delta\alpha\cos\delta_0,\Delta\delta)$ the tangent-plane
offset in radians; this is the limit of K\&S eq.~1 with the
$-\tfrac{1}{2}v_\mathrm{sys}\rho^2$ contraction term (a subtractive
correction for receding systemic velocities) dropped.  The
corrected per-star velocity entering the Gaussian likelihood is
$v_\mathrm{corr} = v_\mathrm{los} - \Delta v_\mathrm{persp}$.

Across the 39-galaxy study sample (LVDB v1.0.5, Pace+ 2022 proper
motions), peak post-selection per-star shifts span $0.04$--$3.72$\,km\,s$^{-1}$
(median ${\sim}0.38$\,km\,s$^{-1}$), with Leo~V the low end and Boötes~III
the high end.  Galaxies whose spectroscopic catalogues extend many
half-light radii beyond the analysis cut (Carina~III: $\rho_\mathrm{cat}/R_h
\simeq 12.7$) show much larger raw shifts on the parent catalogue
(${\sim}6$\,km\,s$^{-1}$ for Carina~III) that collapse to ${<}0.2$\,km\,s$^{-1}$
on the actual Stage~2 sample.  The post-selection magnitudes remain
comparable to or above the per-star velocity uncertainties for the
wider-field galaxies and the high-proper-motion ultra-faints, and
motivate applying the correction throughout the sample.  The dropped quadratic
$\tfrac{1}{2}v_\mathrm{sys}\rho^2$ term remains below
$0.05$\,km\,s$^{-1}$ at $\rho<0.5^\circ$ for $v_\mathrm{sys}\lesssim
300$\,km\,s$^{-1}$, justifying the small-angle approximation; the helper
\code{perspective\_correction\_full} retains the full K\&S expression as
a cross-check.\footnote{Antlia~II is the one exception worth flagging:
with $R_h \simeq 1.7^\circ$ the post-selection sample reaches
$\rho_\mathrm{max} \simeq 2.2^\circ$, where the dropped
$\tfrac{1}{2}v_\mathrm{sys}\rho^2$ term reaches ${\sim}0.21$\,km\,s$^{-1}$
— about $7\%$ of the linear shift's peak ($2.99$\,km\,s$^{-1}$) and the
largest such fraction in the sample.  The correction itself shifts
$\sigma_\mathrm{los}$ by $+0.65$\,km\,s$^{-1}$ ($+10\%$,
${\sim}2\sigma_\mathrm{stat}$); independent verification confirms the
sign — for Antlia~II the per-star perspective shift is anti-correlated
with the velocity residual ($r = -0.43$), so the gradient was
partially cancelling the intrinsic dispersion in $v_\mathrm{los}$ and
removing it broadens the distribution rather than narrowing it.  A
spatial-jackknife test to separate the genuine bulk-PM gradient from
catalog-sampling effects, and a switch to
\code{perspective\_correction\_full}, are left for any future tightening
of precision claims on Antlia~II.}  Ingest now exposes $\mu_{\alpha\ast}$, $\mu_\delta$ and
their asymmetric errors as registry columns (\code{pmra\_mas\_yr},
\code{pmdec\_mas\_yr}, \code{pmra\_em\_mas\_yr}, \dots) and propagates
them to the per-galaxy \code{\_meta}; the compute module is
\code{dwarfjeans.jeans.perspective}.  \code{prepare\_jeans\_input}
subtracts $\Delta v_\mathrm{persp}$ at the LVDB-published $(\mu_{\alpha\ast},
\mu_\delta)$ after the Stage~2 selection cut, stashes the observed
$v_\mathrm{los}$ as \code{V\_observed}, and emits the diagnostics into
\code{audit['perspective']}.  Stage~2 then marginalises over the
proper-motion uncertainty: \code{make\_loglike\_with\_nuisances}
extends the nuisance vector from 7D to 9D by sampling
$(\mu_{\alpha\ast,\mathrm{draw}}, \mu_{\delta,\mathrm{draw}})$ from
split-normal priors built from the LVDB asymmetric errors, and
recomputes $\Delta v_\mathrm{persp}$ per draw using the \emph{sampled}
distance $d$ so that the distance prior propagates correctly through
both the geometric and perspective terms.

A mock-data calibration on an Antlia-II-like configuration ($N_\star =
50$, $d = 124$\,kpc, $R_h \approx 1.4^\circ$, $|\mu| \approx
0.14$\,mas\,yr$^{-1}$; $N = 8$ realisations at \code{nlive}$=100$)
detects no $\sigma_\mathrm{los}(R_h)$ bias from the perspective + PM
pipeline at the standard-error sensitivity the run achieves:
median-of-medians $= 12.246$\,km\,s$^{-1}$ vs truth $12.232$\,km\,s$^{-1}$,
$\mathrm{SE} \approx 0.51$\,km\,s$^{-1}$ ($\approx 4\%$ of truth) from
the realisation-to-realisation dispersion of $1.45$\,km\,s$^{-1}$.
This isolates the correction itself: nuisance priors are pinned at
truth and the PM truth equals the prior central, so this run tests the
perspective + PM \emph{plumbing}, not the full production end-to-end
recovery under realistic nuisance priors or under a data-vs-prior PM
disagreement.  A wider sweep (more realisations, offset PM truth,
production-realistic nuisance priors) is left as a follow-up before
any tightened precision claim.

\subsection{Velocity gradients}
\label{sec:open:gradients}

The Stage~2 likelihood currently assumes a single zero-mean
(systemic-subtracted) dispersion with no spatial trend in
$\langle v_\mathrm{los}\rangle$.  Several classical dwarfs
(e.g.\ Carina, Fornax, Sculptor) show measurable line-of-sight
gradients from a mix of intrinsic rotation, tidal streaming, and
the perspective-motion signal of \S\ref{sec:open:perspective}.
With the perspective correction now in place, the residual
gradient is the rotation/streaming component and the two
contributions must not be double-counted.  Three options are
under consideration --- ingest-time subtraction of a measured
per-galaxy gradient, in-likelihood marginalisation over an
amplitude--position-angle pair, or documenting that the
ultra-faint gradients are consistent with zero and skipping the
correction in that regime.  The implementation is deferred and
tracked alongside the perspective correction.

% ===================================================================
\section*{Acknowledgements}
\todo{Surveys, software, funding.}

\end{document}
